---
title: "Penguins Linear Regression Assignment"
format: html
editor: visual
---

## 1. Load the packages &data

```{r}
options(scipen = 99)
library(tidyverse)
library(bayesrules)
library(rstanarm)
library(ggpubr)
install.packages("ggExtra")
library(ggExtra)
library(report)

data <- bayesrules::penguins_bayes
view(penguins_bayes)
```

## 2. Look before you model

Two quick plots, one sentence each describing what you see:

- A histogram of `flipper_length_mm`. Where is it centered? Roughly bell-shaped, or doing something unusual?
  - It is centered at 190 with some concentration near 220. It is roughly bell-shaped with two bells.
- A scatterplot of `flipper_length_mm` (y) against `bill_length_mm` (x). Linear? Positive or negative? Strong or weak? It is mostly linear and positive.

```{r}
ggplot(data = penguins_bayes, aes(x = flipper_length_mm)) +
  geom_histogram()
```

```{r}
ggplot(data = penguins_bayes, aes(x = flipper_length_mm, y = bill_length_mm)) +
  geom_point() +
  geom_smooth(method = "lm") +
  stat_cor() 
```

```{r}
p <- ggplot(data = penguins_bayes, aes(x = flipper_length_mm, y = bill_length_mm)) +
  geom_point() +
  theme_bw()

# Add the side marginal plots (options: "density", "histogram", or "boxplot")
ggMarginal(p, type = "density")  
```

## 3. Translate these three beliefs into priors

Read each statement carefully — you need to convert plain language into specific numbers.

> **Typical flipper length.** Across all three species in this dataset, a reasonable guess for average flipper length is around 200mm — with a plausible range of roughly 180mm to 220mm.

> **The bill-to-flipper relationship.** Bill length and flipper length are both rough proxies for body size, so for every 1mm increase in bill length, flipper length typically increases by around 3.5mm — with a plausible range of roughly 1.5mm to 5.5mm.

> **Leftover variability.** Even penguins with identical bill lengths won't have identical flippers. A reasonable guess for the typical leftover spread is around 12.5mm.

**How to convert each one:**

- For a Normal prior, "around X, plausibly between A and B" means: X is your mean, and the width from A to B is your **95% range**, which equals **mean ± 1.96 × SD**. Solve for SD.
- For sigma's Exponential prior, there's no range to convert — just a single target mean. Recall **E(sigma) = 1/rate**, so `rate = 1/E(sigma)`.

```{r}
plot_normal(200,10.2)
```

```{r}
plot_normal(3.5, 1.02)
```

```{r}
plot_gamma(1,0.08)
```

```{r}
m0 <- 200   # centered intercept mean
s0 <- 10.2   # centered intercept SD, half range of the confidence interval(220-180) / 2 = 20 / 1.98 = 10.2
m1 <- 3.5   # slope mean
s1 <- 1.02   # slope SD
l  <- 0.08   # sigma's exponential rate, 1 / 12.5 = 0.08
```

Write your five numbers as plain variables before moving on:

``` r
m0 <- 200   # centered intercept mean
s0 <- 10.2   # centered intercept SD, half range of the confidence interval(220-180) / 2 = 20 / 1.98 = 10.2
m1 <- 3.5   # slope mean
s1 <- 1.02   # slope SD
l  <- 0.08   # sigma's exponential rate, 1 / 12.5 = 0.08
```

## 4. Fit the model

```{r}
model <- stan_glm(
  flipper_length_mm ~ bill_length_mm, data = data,
  family = gaussian,
  prior_intercept = normal(m0, s0),
  prior = normal(m1, s1),
  prior_aux = exponential(l),
  chains = 4, iter = 2000, seed = 84735
)
```

```{r}
model
```

```{r}
summary(model)
```

```{r}
plot_normal(1.7, 0.051)
```

```{r}
report::report(model)
```

<https://utoronto.scholaris.ca/server/api/core/bitstreams/20be7013-0cf4-4de3-81ef-7713db5861cc/content>

<https://www.sfu.ca/~tswartz/papers/tennis1.pdf>

**Before reading your results, a critical note:** the `(Intercept)` row in your model's printout is *not* the same number as `m0`. `m0`/`s0` describe **beta0c**, the typical flipper length **at the average bill length** in the dataset. `stan_glm()` always reports the **raw** intercept instead — typical flipper length at a bill length of *zero*, which is not a meaningful scenario and will look nothing like `m0`. To get the number you actually tuned a prior for, you have to compute it yourself from the posterior draws:

```{r}
posterior <- as.data.frame(model)
beta0c_draws <- posterior$`(Intercept)` + posterior$bill_length_mm * mean(data$bill_length_mm)
```

`beta0c_draws` is now a vector of posterior draws for the *centered* intercept — use this, not the raw `(Intercept)` column, for anything below that asks about "typical flipper length at the average bill length."

## 5. Report your results

For each item below, the exact function to use is named — plug in the right column name and run it.

**1. The posterior median relationship.**

```{r}
median(posterior$`(Intercept)`)
median(posterior$bill_length_mm)
```

Write the fitted equation: flipper_length_mm = \_\_\_ + \_\_\_ × bill_length_mm.

**2. A 90% credible interval for the slope.**

```{r}
quantile(posterior$bill_length_mm, c(0.05, 0.95))
```

One sentence: what does this range say about how flipper length changes with bill length?

**3. A 90% credible interval for the centered intercept.**

```{r}
quantile(beta0c_draws, c(0.05, 0.95))
```

Use `beta0c_draws` here, not the raw `(Intercept)` column — see the note in Step 4.

**4. The posterior probability the slope is positive.**

```{r}
mean(posterior$bill_length_mm > 0)
```

One sentence: what does this number tell you that a p-value wouldn't?

**5. MCMC diagnostics.**

```{r}
summary(model)[, c("Rhat", "n_eff")]
```

Confirm Rhat is close to 1 for every parameter before trusting anything above.

**6. Compare to a plain linear model.**

```{r}
lm(flipper_length_mm ~ bill_length_mm, data = data)
```
